%% doc.tex

\input{prelude.tex}

\begin{document}

\title{too old to demonstrate, to young to die}
\author{Marcel Taude}
\date{November 2015}

\begin{abstract}
	Playing with integer equations is really nice. This paper is a ... hmmm ... demonstration of an old adage \emph{if you don't practice mathematics for a long time, it's hard to come back to it}. Even when listening cool jazz from \emph{Lee Konitz}. At least, I played with \LaTeX\ and \emph{gcc}.

	At first, I wanted to find an algorithm to find integers which verify $a^3 + b^3 + c^3 = 1000^2.a + 1000.b + c$ with $a, b, c \in [0, 1000[$. There is an obvious lazy algorithm which tries every values of $a$, $b$ and $c$ in the range $[0, 1000[$. I wanted a clever one, not sure I found it, but read.
\end{abstract}	

\maketitle

\tableofcontents

\lstset{basicstyle=\small\ttfamily,language=make,morekeywords={ifeq,ifneq,ifdef,ifndef,else,endif,addprefix,notdir,PHONY}, keywordstyle=\bfseries\color{kword},backgroundcolor=\color{lgray},breaklines=true}


\section{the problem}
Find all integers which verify :
\begin{equation}\label{eq:start}
	a^3 + b^3 + c^3 = 1000^2.a + 1000.b + c \text{, with } a, b, c \in [0, 1000[
\end{equation}
We can \emph{easily} find all the integers $(a, b, c)$ verfiying \eqref{eq:start} by computing the two members of \eqref{eq:start} for every values $(a, b, c)$ from 0 to 999. Can we find a more clever way to get these values?

We can transform \eqref{eq:start} in :
\begin{equation}\label{eq:second}
\begin{split}
	c^3 - c &= 1000^2.a + 1000.b - a^3 - b^3 \\
	c(c - 1)(c + 1) &= a(1000 - a)(1000 + a) + b(1000 - b^2)
 \end{split}
\end{equation}

It becomes more obvious that the triplets $(0, 0, 0)$ and $(0, 0, 1)$ are two solutions of \eqref{eq:start}.

\section{... and this is the place where I became mad}
So much time without playing with mathematics makes things harder than I thought. Music helps me to keep cool, some good old jazz played by \emph{Lee Konitz} in the late fifties, and that's all. At least, I am playing with \LaTeX and \emph{gcc}...

\section{reboot}
We will do it in a lighter way. We started with power 3, bad idea. 

\begin{equation}\label{eq:pow2}
	a^2 + b^2 = 100.a + b \text{ with } a, b \in [0, 100[
\end{equation}

We can make this transformation :
\begin{equation}\label{eq:pow2bis}
	b(b-1) = a(100 - a)
\end{equation}

We can see that the left side of \eqref{eq:pow2ter} must be even, so $a$ is even too.

\subsection{the even side}
We begin with the case of $b$ even :
\begin{equation}\label{eq:pow2ter}
\begin{split}
	a &= 2 \alpha \\
	b &= 2 \beta \\
	2 \beta(2 \beta - 1) &= 2 \alpha(100 - 2 \alpha) \\
	\beta(2\beta - 1) &= 2 \alpha(50 - \alpha)
\end{split}
\end{equation}

If so, $\beta$ must be even, \emph{i.e.} :
\begin{equation}\label{eq:pow2:4}
\begin{split}
	\beta &= 2 \gamma \\
	2 \gamma (4 \gamma - 1) &= 2 \alpha (50 - \alpha) \\
	\gamma (4 \gamma - 1) &= \alpha (50 - \alpha)
\end{split}
\end{equation}

The good news is : 
\begin{equation}\label{eq:pow2:5}
\begin{split}
	\gamma &\in [0, 25[ \\
	\alpha &\in [0, 50[ \\
	a &= 2 \alpha \\
	b &= 4 \gamma
\end{split}
\end{equation}

We have less couples $(\alpha, \gamma)$ to test.

\subsection{the odd side}
Now, we are looking at odd values of $b$:
\begin{equation}\label{eq:pow2odd}
\begin{split}
	a &= 2 \alpha \\
	b &= 2 \beta + 1\\
	2 \beta(2 \beta + 1) &= 2 \alpha(100 - 2 \alpha) \\
	\beta(2\beta + 1) &= 2 \alpha(50 - \alpha)
\end{split}
\end{equation}

We comme back to the previous case with few changes :
\begin{equation}\label{eq:pow2:5}
\begin{split}
	\gamma &\in [0, 25] \\
	\alpha &\in [0, 50[ \\
	a &= 2 \alpha \\
	b &= 4 \gamma + 1
\end{split}
\end{equation}

%% \subsection{conclusion}



\end{document}
